Type Layers for Gradual Verification -- Industrial Internship Report

### 2023-03-24 tex/report_uds/report.tex
* Generalized liquid layer with layer identifier
Added missing graph

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Industry Internship -- Final Report}

@@

\titlerunning{Industrial Internship Report}

@@

\begin{tabularx}{\textwidth}{rX}
	\textbf{Name of Intern/Student:} & Lukas Abelt\\
	\textbf{Student Contact Information:} & \email{luab00001@stud.uni-saarland.de}\\
	\textbf{Matriculation Number at UdS:} & 7009048\\
	\textbf{Course of Study at UdS:} & Master of Science Computer Science\\
	\textbf{Start of Internship:} & 01.01.2022\\
	\textbf{End of Internship:} & 31.03.2023\\
	\textbf{Hosting Insitution:} & LASIGE Research Unit\footnote{\url{https://www.lasige.pt/}}\newline Reliable Software Systems Research Line\newline Faculdade de Ciências da Universidade de Lisboa\\
	\textbf{Internship Supervisor:} & Alcides Fonseca, Assistant Professor\\
	\textbf{Supervisor Contact Information:} & \email{amfonseca@fc.ul.pt}\\
	\textbf{Internship Title:} & \textit{Type Layers for Gradual Verification}
\end{tabularx}

\section{Internship Description}
This section will give a bread overview of the work that was carried out during the internship. It wstart by providing context to the problem at hand and continue further with the proposed approach and goals of the internship. A technical description of the implemented solution will be provided and finally a short overview of the acquired skills and produced deliverables will be provided.

@@

The prototype consists of a simple compiler and interpreter for said functional language. The prototype has been developed in Python using the \Lark compiler framework. The core of our compiler is a \textit{layered static verifier}. Within a program, the programer can define annotations to be passed on to one or multiple verification layers. The implementation of each verification layer is provided by the programer in a separate python file with a pre-defined set of functions that will be called by the verifier. Specifically, each layer can provide the following information:

\begin{figure}[ht!]
	\centering
	\includegraphics[width=.8\textwidth]{framework_overview}
	\label{fig:framework_overview}
	\caption{Conceptual Overview of \LayeredTypes}
\end{figure}

In the second Use-Case we focus on a problem currently present in implementations for \textit{Liquid Types}:

\subsection{Deliverables and Publications}

\bibliographystyle{plain}
\bibliography{literature.bib}
\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[caption={Simple example code},label={lst:code_before}]
printLines lines idx len {
  if len != idx then {
    line = get(lines, idx)
    print(line)
    printLines(lines, idx+1, len)
  }
}

@@

>>>>

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Type Layers for Gradual Verification -- Industrial Internship Report}

@@

\titlerunning{Type Layers for Gradual Verification}

@@

\begin{tabularx}{\textwidth}{rX}
	\textbf{Name of Intern/Student:} & Lukas Abelt\\
	\textbf{Student Contact Information:} & \email{luab00001@stud.uni-saarland.de}\\
	\textbf{Matriculation Number at UdS:} & 7009048\\
	\textbf{Course of Study at UdS:} & Master of Science Computer Science\\
	\textbf{Start of Internship:} & 01.01.2022\\
	\textbf{End of Internship:} & 31.03.2023\\
	\textbf{Hosting Insitution:} & LASIGE Research Unit\footnote{\url{https://www.lasige.pt/}}\newline Reliable Software Systems Research Line\newline Faculdade de Ciências da Universidade de Lisboa\\
	\textbf{Working hours per week} & 35 \\
	\textbf{Internship Supervisor:} & Alcides Fonseca, Assistant Professor\\
	\textbf{Supervisor Contact Information:} & \email{amfonseca@fc.ul.pt}\\
	\textbf{Internship Title:} & \textit{Type Layers for Gradual Verification}\\ 
\end{tabularx}
\section{Signatures}

I hereby acknowledge that the information provided in this report is accurate and sufficiently describes the work carried out.
\vfill
\par\noindent\makebox[2.5in]{\hrulefill} \hfill\makebox[2.5in]{\hrulefill}%
\par\noindent\makebox[2.5in][l]{Lukas Abelt, Intern}      \hfill\makebox[2.5in][l]{Alcides Fonseca, Supervisor}%
\par\noindent\makebox[2.5in][l]{Lisbon, \today}      \hfill\makebox[2.5in][l]{Lisbon, \today}%
\newpage
\section{Internship Description}
This section will give a bread overview of the work that was carried out during the internship. It wstart by providing context to the problem at hand and continue further with the proposed approach and goals of the internship. A technical description of the implemented solution will be provided and finally a short overview of the acquired skills and produced deliverables will be provided.

@@

\begin{minipage}{0.4\linewidth}
\begin{lstlisting}[caption={Simple example code},label={lst:code_before}]
printLines lines idx len {
  if len != idx then {
    line = get(lines, idx)
    print(line)
    printLines(lines, idx+1, len)
  }
}

@@

The prototype consists of a simple compiler and interpreter for said functional language. The prototype has been developed in Python using the \Lark compiler framework. The core of our compiler is a \textit{layered static verifier}. Within a program, the programer can define annotations to be passed on to one or multiple verification layers. An example of such annotations is shown in \Cref{lst:code_after} The implementation of each verification layer is provided by the programer in a separate python file with a pre-defined set of functions that will be called by the verifier. Specifically, each layer can provide the following information:

\begin{figure}[ht!]
	\centering
	\includegraphics[width=.8\textwidth]{framework_overview}
	\caption{Conceptual Overview of \LayeredTypes}
	\label{fig:framework_overview}
\end{figure}

In the second Use-Case we focus on a problem currently present in implementations for \textit{Liquid Types}. When operating on liquid types, often we want to reason about orthogonal properties of variables and functions. One example could be the verification of machine learning dataset pipelines: Some operations may only depend and influence the rows (Datapoints) of a dataset (e.g. adding a datapoint does not affect the number of columns), while others only only affect the columns (Features). Finally, some actions require information on both rows and columns. For example when training a model based on a dataset, we may want to ensure that this dataset contains enough datapoints for the amount of features.

In our specific running example we added a few simple functions with that specific behavior. We have several methods (Adding, removing, extracting a certain amount of elements) that only affect rows and columns separately but also a \texttt{train} method, that introduces a relationship between the number of columns and the number of rows: We assume that we need at least ten times as many datapoints as we have features.

In our example we showcased the issues that arise with hard to interpret error messages that occur when typechecking fails when verifying all properties at once. Subsequently we separated the individual properties and interactions into separate layers to demonstrate how this separation affects the error messages produced.

\subsection{Deliverables, Publications and Skills acquired}

During the work on this internship and the subsequent implementation, the intern acquired theoretical as well as practical knowledge on the following aspects:
\begin{itemize}
	\item Compiler Construction
	\item Verification Techniques
	\item Importance and Distinction of different Type Systems including, but not limited to:
		\begin{itemize}
			\item Linear Types
			\item Dependent Types
			\item Liquid Types
		\end{itemize}
\end{itemize}

\bibliographystyle{plain}
\bibliography{literature.bib}
